### 2019-08-30 paper/reply_to_referee.txt
Start on a reply to the referee of the mixed hybrid paper.

----------------------------------------------------------------------
Reviewer's Comments:

> This paper focuses on the outcome of thermonuclear explosions of near
> Chandrasekhar mass white dwarfs coming from the narrow range of IMS that
> produce hybrid C+O+Ne cores. The purpose of the work is to determine
> the differences between the thermonuclear explosion of mixed vs unmixed
> hybrid cores, motivated by the recent finding that the cores undergo a
> phase of convective mixing that destroys the differentiated structure
> made of a C+O nut surrounded by O+Ne layers. The approach consists on
> running the MESA code along the pre-explosive phase, followed by two
> dimensional hydro simulations of the explosion. There are also calculated
> the explosion of pure C+O white dwarfs starting from identical central
> density and temperature as the C+O+Ne mixed white dwarfs. The conclusions
> are that the mixed structures lead to lower yields of 56Ni as compared
> to the pure C+O case as well as to the unmixed hybrid cores.

> The paper is timely and interesting, and it merits publication in ApJ
> after some points are addressed, and the answers be reflected in the
> text. Some of these points are of methodological character, while others
> are more general.

> *Methodology*

> - It is not clear to me if the input deflagration speed (pg 7) has been
> recalibrated to the particular initial composition of the mixed hybrid
> core. Willcox et al (2016) just reports recomputing detonations, but there
> is no mention of deflagrations (at least I have not been able to find it).

> - Why is the MESA calculation stopped and mapped to FLASH just when
> the central temperature reaches log Tc = 8.9? What is the criterion for
> choosing such a temperature? Does it depend on composition?

> - Why was the C+O model constructed with the same central temperature and
> density as the mixed hybrid model? Would it be more appropriate to follow
> a C+O white dwarf with MESA until some runaway condition was accomplished?

> - Related to the last point above: it is not clear how was assigned the
> composition profile to the C+O model. In particular, why is the abundance
> of 22Ne in the C+O model smaller than that in the hybrid mixed model? Does
> it has an impact on the final yield of 56Ni?

> - Why were the models calculated in this paper ignited at the center,
> unlike the models in Willcox et al (2016)? Note that you invoke the
> off-center ignition as one of the causes of the differing results, but you
> do not provide a justification of the different ignition configuration.

> - In the first paragraph of section 3, it is argued that the yield of 56Ni
> was approximated by assuming the composition upon NSE freeze-out is 56Ni
> plus equal parts 54Fe and 58Ni. However, in many models of supernova Ia
> there is a large abundance of 56Ni in the layers that do not go beyond
> QNSE. Have you corrected for this effect somehow?

> *General*

> - Which is the role played by the hybrid progenitors in the global plot of
> supernovae type Ia? Since the lower limit of the 56Ni mass distribution
> is 0.6 solar masses, does it mean that hybrid progenitors lead to
> normal-to-bright supernovae while C+O progenitors represent the brighter
> end of this kind of supernovae? Does it imply a dominant contribution
> of double degenerate channel for the normal supernovae Ia? Most readers
> would thank a brief discussion of these and related issues.

> - Please, provide the values of the carbon abundance at the center and
> the mean of it in the white dwarf, in the initial model as well as in
> the final model of the MESA calculation.

> - Subsections 1.1 and 1.2 are more typical of a review than a regular
> paper, and are not at all necessary to understand what follows. They
> should be shortened to one or two medium length paragraphs. By the way,
> the reference Nomoto et al (1984) does not match the description of the
> flame that precedes it.

> - Figures 3, 6, and 7 are not quite informative, and the same applies
> to the short mention of them in the text. The authors should consider
> removing them.