### 2019-08-31 paper/reply_to_referee.txt
More work on the new hybrid model manuscript to address the referee's concerns.

----------------------------------------------------------------------
Reviewer's Comments:

> This paper focuses on the outcome of thermonuclear explosions of near
> Chandrasekhar mass white dwarfs coming from the narrow range of IMS that
> produce hybrid C+O+Ne cores. The purpose of the work is to determine
> the differences between the thermonuclear explosion of mixed vs unmixed
> hybrid cores, motivated by the recent finding that the cores undergo a
> phase of convective mixing that destroys the differentiated structure
> made of a C+O nut surrounded by O+Ne layers. The approach consists on
> running the MESA code along the pre-explosive phase, followed by two
> dimensional hydro simulations of the explosion. There are also calculated
> the explosion of pure C+O white dwarfs starting from identical central
> density and temperature as the C+O+Ne mixed white dwarfs. The conclusions
> are that the mixed structures lead to lower yields of 56Ni as compared
> to the pure C+O case as well as to the unmixed hybrid cores.

> The paper is timely and interesting, and it merits publication in ApJ
> after some points are addressed, and the answers be reflected in the
> text. Some of these points are of methodological character, while others
> are more general.

We thank the referee for reading out manuscript and offering very
constructive criticism and suggestion. We are in almost complete agreement
with the referee and have modified the manuscript accordingly. We address
our changes for each point below. As a result of our revisions, we believe 
the manuscript is stronger and we hope it is now ready for publication.

> *Methodology*

> - It is not clear to me if the input deflagration speed (pg 7) has been
> recalibrated to the particular initial composition of the mixed hybrid
> core. Willcox et al (2016) just reports recomputing detonations, but there
> is no mention of deflagrations (at least I have not been able to find it).

We regret no explaining the methodology of our deflagration speeds. The 
answer is that the deflagration speeds for input to the flame model are 
from tabulated results. Laminar flame speeds are first 
obtained by interpolation within a
three-dimensional table of 12C, 22Ne, and log density obtained from
prior direct numerical simulations. Here 22Ne is uses as a proxy for
for neutron-rich elements so the interpolated laminar flame speeds capture
the boost due to neutronization. These laminar flame speed are
then boosted to account for the effects of unresolved buoyancy
and background turbulence and passed to the flame model. We have
updated the text of the manuscript to explain the flame speeds
and included a citation to earlier work describing the tabulated
flame speeds in detail.
[NOTE: CHECK THIS. The models include 20Ne. Does the above
description properly account for that? As I recall, nothing new
was done to the tables for the hybrid Willcox et al. models.]

Also, the detonation tests in Willcox et al. were tests of the
energy release of the flame model and not flame speed tests. The 
model identical amounts of energy for both deflagrations and 
detonations in identical material, so detonation simulations were
a simple check of the modifications we made to the model to
include 20Ne in the mix.

> - Why is the MESA calculation stopped and mapped to FLASH just when
> the central temperature reaches log Tc = 8.9? What is the criterion for
> choosing such a temperature? Does it depend on composition?

Again, we did not explain our methodology well. This temperature is
very close to that required to ignite the deflagration of the
first phase of the explosion, so the progenitor is in a ready 
for ignition state. Yes, it is a (somewhat mildly) a function of 
composition, but at this temperature these models have very little 
evolution left before the flame would be born even with differing 
composition. We have updated the text to better explain this issue.

> - Why was the C+O model constructed with the same central temperature and
> density as the mixed hybrid model? Would it be more appropriate to follow
> a C+O white dwarf with MESA until some runaway condition was accomplished?

> - Related to the last point above: it is not clear how was assigned the
> composition profile to the C+O model. In particular, why is the abundance
> of 22Ne in the C+O model smaller than that in the hybrid mixed model? Does
> it has an impact on the final yield of 56Ni?

The referee raises an interesting point about how one might perform
these parameter studies. Indeed, a more realistic C/O progenitor as
would be produced by MESA would be "state of the art." The parameterized
C/O models we chose are motivated by results from stellar evolution and,
in particular, have compositions that reflect the effect of ``simmering", 
that is slow convective burning, in the late stages of evolution as the
WD approaches the Chandrasekhar limit. This effect is observed as
the lower C fraction interior to about r =1.2e8 cm (the solid curve
in the bottom panel of Figure 1). Accordingly, while parameterized, 
the structure of C/O progenitor does reflect our current understanding 
of the state just prior to ignition.

For this study we decided to use these parameterized C-O models because
being able to construct them to match the central density and 
temperature of the hybrid models allowed us to better control the 
differences between the hybrid and traditional models. More realistic
C-O models would not match the conditions of the hybrid models, and
as we have shown in previous work, differences in density can have
a substantial effect on the yields.

We have added text to the manuscript and slightly re-arranged the 
description to better explain the motivation for our choice.

> - Why were the models calculated in this paper ignited at the center,
> unlike the models in Willcox et al (2016)? Note that you invoke the
> off-center ignition as one of the causes of the differing results, but you
> do not provide a justification of the different ignition configuration.

> - In the first paragraph of section 3, it is argued that the yield of 56Ni
> was approximated by assuming the composition upon NSE freeze-out is 56Ni
> plus equal parts 54Fe and 58Ni. However, in many models of supernova Ia
> there is a large abundance of 56Ni in the layers that do not go beyond
> QNSE. Have you corrected for this effect somehow?

> *General*

> - Which is the role played by the hybrid progenitors in the global plot of
> supernovae type Ia? Since the lower limit of the 56Ni mass distribution
> is 0.6 solar masses, does it mean that hybrid progenitors lead to
> normal-to-bright supernovae while C+O progenitors represent the brighter
> end of this kind of supernovae? Does it imply a dominant contribution
> of double degenerate channel for the normal supernovae Ia? Most readers
> would thank a brief discussion of these and related issues.

> - Please, provide the values of the carbon abundance at the center and
> the mean of it in the white dwarf, in the initial model as well as in
> the final model of the MESA calculation.

> - Subsections 1.1 and 1.2 are more typical of a review than a regular
> paper, and are not at all necessary to understand what follows. They
> should be shortened to one or two medium length paragraphs. By the way,
> the reference Nomoto et al (1984) does not match the description of the
> flame that precedes it.

> - Figures 3, 6, and 7 are not quite informative, and the same applies
> to the short mention of them in the text. The authors should consider
> removing them.