### 2019-09-23 paper/reply_to_referee.txt
Updated one of the questions based on a conversation with Don.

----------------------------------------------------------------------
Reviewer's Comments:

> This paper focuses on the outcome of thermonuclear explosions of near
> Chandrasekhar mass white dwarfs coming from the narrow range of IMS that
> produce hybrid C+O+Ne cores. The purpose of the work is to determine
> the differences between the thermonuclear explosion of mixed vs unmixed
> hybrid cores, motivated by the recent finding that the cores undergo a
> phase of convective mixing that destroys the differentiated structure
> made of a C+O nut surrounded by O+Ne layers. The approach consists on
> running the MESA code along the pre-explosive phase, followed by two
> dimensional hydro simulations of the explosion. There are also calculated
> the explosion of pure C+O white dwarfs starting from identical central
> density and temperature as the C+O+Ne mixed white dwarfs. The conclusions
> are that the mixed structures lead to lower yields of 56Ni as compared
> to the pure C+O case as well as to the unmixed hybrid cores.

> The paper is timely and interesting, and it merits publication in ApJ
> after some points are addressed, and the answers be reflected in the
> text. Some of these points are of methodological character, while others
> are more general.

We thank the referee for reading our manuscript and offering very
constructive criticism and suggestions. We are in almost complete agreement
with the referee and have modified the manuscript accordingly. We address
our changes for each point below. As a result of our revisions, we believe 
the manuscript is stronger and we hope it is now ready for publication.

> *Methodology*

> - It is not clear to me if the input deflagration speed (pg 7) has been
> recalibrated to the particular initial composition of the mixed hybrid
> core. Willcox et al (2016) just reports recomputing detonations, but there
> is no mention of deflagrations (at least I have not been able to find it).

We regret no explaining the methodology of our deflagration speeds. The 
answer is that the deflagration speeds for input to the flame model are 
from tabulated results. Laminar flame speeds are first 
obtained by interpolation within a
three-dimensional table of 12C, 22Ne, and log density obtained from
prior direct numerical simulations. Here 22Ne is uses as a proxy for
for neutron-rich elements so the interpolated laminar flame speeds capture
the boost due to neutronization. These laminar flame speed are
then boosted to account for the effects of unresolved buoyancy
and background turbulence and passed to the flame model. We have
updated the text of the manuscript to explain the flame speeds
and included a citation to earlier work describing the tabulated
flame speeds in detail.

[AC NOTE: CHECK THE ABOVE. The models include 20Ne. Does the above
description properly account for that? As I recall, nothing new
was done to the tables for the hybrid Willcox et al. models.]

[rephrasing the question- does 20Ne count as fuel for both
the energy release and the flame speed from the table? Don seems 
to recall that we had the energy release of burning 20Ne in the 
detonation tests, which says it was included for the energy 
release. But do we just consider 12C + 12C buring for the
flame speed?]

Also, the detonation tests in Willcox et al. were tests of the
energy release of the flame model and not flame speed tests. The 
model releases identical amounts of energy for both deflagrations and 
detonations in identical material, so detonation simulations were
a simple check of the modifications we made to the model to
include 20Ne in the mix.

> - Why is the MESA calculation stopped and mapped to FLASH just when
> the central temperature reaches log Tc = 8.9? What is the criterion for
> choosing such a temperature? Does it depend on composition?

Again, we did not explain our methodology well. This temperature is
very close to that required to ignite the deflagration of the
first phase of the explosion, so the progenitor is in a ready 
for ignition state. Yes, it is a (somewhat mildly) a function of 
composition, but at this temperature these models have very little 
evolution left before the flame would be born even with differing 
composition. We have updated the text to better explain this issue.

> - Why was the C+O model constructed with the same central temperature and
> density as the mixed hybrid model? Would it be more appropriate to follow
> a C+O white dwarf with MESA until some runaway condition was accomplished?

> - Related to the last point above: it is not clear how was assigned the
> composition profile to the C+O model. In particular, why is the abundance
> of 22Ne in the C+O model smaller than that in the hybrid mixed model? Does
> it has an impact on the final yield of 56Ni?

The referee raises an interesting point about how one might perform
these parameter studies. Indeed, a more realistic C/O progenitor as
would be produced by MESA would be "state of the art." The parameterized
C/O models we chose are motivated by results from stellar evolution and,
in particular, have compositions that reflect the effect of ``simmering", 
that is slow convective burning, in the late stages of evolution as the
WD approaches the Chandrasekhar limit. This effect is observed as
the lower C fraction interior to about r =1.2e8 cm (the solid curve
in the bottom panel of Figure 1). Accordingly, while parameterized, 
the structure of C/O progenitor does reflect our current understanding 
of the state just prior to ignition.

For this study we decided to use these parameterized C-O models because
being able to construct them to match the central density and 
temperature of the hybrid models allowed us to better control the 
differences between the hybrid and traditional models. More realistic
C-O models would not match the conditions of the hybrid models, and
as we have shown in previous work, differences in density can have
a substantial effect on the yields.

We have added text to the manuscript and slightly re-arranged the 
description to better explain the motivation for our choice.

> - Why were the models calculated in this paper ignited at the center,
> unlike the models in Willcox et al (2016)? Note that you invoke the
> off-center ignition as one of the causes of the differing results, but you
> do not provide a justification of the different ignition configuration.

Something else we could have better explained. The hybrid models in Willcox
et al have a very different structure with the highest temperature, where
ignition of the deflagration is expected, at a radius of about 300 km. 
The profiles are illustrated in Figures 2 and 9 of Willcox et al. Yet 
another thing we have made clear in the revised manuscript.

> - In the first paragraph of section 3, it is argued that the yield of 56Ni
> was approximated by assuming the composition upon NSE freeze-out is 56Ni
> plus equal parts 54Fe and 58Ni. However, in many models of supernova Ia
> there is a large abundance of 56Ni in the layers that do not go beyond
> QNSE. Have you corrected for this effect somehow?

[AC: Hmmm. I think this is not right. What to say?]

The estimate of 56Ni yield does follow from assuming the NSE mix is
composed of 56Ni plus equal parts 54Fe and 58Ni, and this approach is
how all 56Ni in the simulation is calculated. 56Ni in (outer) layers
that did not go beyond QNSE would have been synthesized by fusion 
going to the Fe-group and then mixed. As such, the estimate we use 
should be robust. We have added a sentence with citations and an
additional citation to previous work verifying this approach in a variety 
of scenaros to the paragraph.

> *General*

> - Which is the role played by the hybrid progenitors in the global plot of
> supernovae type Ia? Since the lower limit of the 56Ni mass distribution
> is 0.6 solar masses, does it mean that hybrid progenitors lead to
> normal-to-bright supernovae while C+O progenitors represent the brighter
> end of this kind of supernovae? Does it imply a dominant contribution
> of double degenerate channel for the normal supernovae Ia? Most readers
> would thank a brief discussion of these and related issues.

These are the questions that we are exploring with this and related work,
and we regret not make these clear in the introductory material. We
have added text addressing exactly these issues.

> - Please, provide the values of the carbon abundance at the center and
> the mean of it in the white dwarf, in the initial model as well as in
> the final model of the MESA calculation.

The central carbon abundance in the MESA model is 0.1419 and the initial
model for Flash is constructed to have the same central carbon abundance.
The central carbon abundance in the C-O model is 0.4. We have added these
numbers to the manuscript. The mean values of C abundances were not
calculated with the models. We could integrate the models
to obtain a mass-weighted average, but we are not sure of the utility
of this quantity as our results show that the yields are a sensitive
function of the density. We prefer to just state the central C abundances
and present the profiles as we did in Figure 1 and hope this is acceptable.

> - Subsections 1.1 and 1.2 are more typical of a review than a regular
> paper, and are not at all necessary to understand what follows. They
> should be shortened to one or two medium length paragraphs. By the way,
> the reference Nomoto et al (1984) does not match the description of the
> flame that precedes it.

We have substantially shortened the sections, including removing the
Nomoto (1984) reference (good catch!). We modified both sections
with an emphasis on just the points important to our analysis (e.g. 
the duration of the deflagration phase and expansion).

> - Figures 3, 6, and 7 are not quite informative, and the same applies
> to the short mention of them in the text. The authors should consider
> removing them.

We appreciate the referee's point. The figures show the evolution of quantities 
like 56Ni abundance with time and are similar to some of the figures showing 
abundance evolution in Willcox et al. Because they allow the reader to both 
visualize the scatter in the yields and also the character of the burning 
(such as the DDT), we feel they are useful and would like to keep them. Also,
they are similar to figures in Willcox et al, which allows comparison
to results from another class of hybrid progenitors. We agree with the referee 
that the discussion was scant and have improved that.

[AC: CHECK RED TEXT IN THE CAPTION OF FIG. 6 and similar text in the description]

In summary. we have taken the referee's criticism and suggestions and revised 
the manuscript accordingly. The result is a significantly improved version we trust 
is ready for publication.