### 2019-10-25 paper/reply_to_referee.txt

----------------------------------------------------------------------
Reviewer's Comments:

> This paper focuses on the outcome of thermonuclear explosions of near
> Chandrasekhar mass white dwarfs coming from the narrow range of IMS that
> produce hybrid C+O+Ne cores. The purpose of the work is to determine
> the differences between the thermonuclear explosion of mixed vs unmixed
> hybrid cores, motivated by the recent finding that the cores undergo a
> phase of convective mixing that destroys the differentiated structure
> made of a C+O nut surrounded by O+Ne layers. The approach consists on
> running the MESA code along the pre-explosive phase, followed by two
> dimensional hydro simulations of the explosion. There are also calculated
> the explosion of pure C+O white dwarfs starting from identical central
> density and temperature as the C+O+Ne mixed white dwarfs. The conclusions
> are that the mixed structures lead to lower yields of 56Ni as compared
> to the pure C+O case as well as to the unmixed hybrid cores.

> The paper is timely and interesting, and it merits publication in ApJ
> after some points are addressed, and the answers be reflected in the
> text. Some of these points are of methodological character, while others
> are more general.

We thank the referee for reading our manuscript and offering
very constructive criticism and suggestions. We are in almost
complete agreement with the referee and have modified the manuscript
accordingly. We address our changes for each point below, and the changes
we made to the manuscript are in blue in the revised version. As a result
of our revisions, we believe the manuscript is stronger and we hope it
is now ready for publication.

> *Methodology*

> - It is not clear to me if the input deflagration speed (pg 7) has been
> recalibrated to the particular initial composition of the mixed hybrid
> core. Willcox et al (2016) just reports recomputing detonations, but there
> is no mention of deflagrations (at least I have not been able to find it).

We regret not explaining the methodology of our deflagration speeds. The
answer is that the deflagration speeds for input to the flame model
are from tabulated results. Laminar flame speeds are first obtained by
interpolation within a three-dimensional table of 12C, 22Ne, and log
density obtained from prior direct numerical simulations. Here 22Ne is
uses as a proxy for for neutron-rich elements so the interpolated laminar
flame speeds capture the boost due to neutronization. We use the same
tables that were used in Willcox et al. and previous work. These laminar
flame speed are then boosted to account for the effects of unresolved
buoyancy and background turbulence and passed to the flame model. We have
updated the text of section 2.1 to explain the flame speeds and included a
citation to earlier work describing the tabulated flame speeds in detail.

Also, the detonation tests in Willcox et al. were tests that the
addition of 20Ne did not alter the four-stage structure assumed by
the flame model. The ZND detonations indicated that the burning still
proceeds through four main stages of burning distinguished by "steps"
in the density/pressure profiles behind the ZND shock corresponding to
mass fraction crossing times for fuel->ash, ash->NSQE, NSQE->NSE. This
justified the minimal modification to the burning in FLASH where
we included 20Ne in the initial composition and included its energy
generation assuming 20Ne burns to 24Mg.

> - Why is the MESA calculation stopped and mapped to FLASH just when
> the central temperature reaches log Tc = 8.9? What is the criterion for
> choosing such a temperature? Does it depend on composition?

Again, we did not explain our methodology well. This temperature is
very close to that required to ignite the deflagration of the first
phase of the explosion, so the progenitor is in a ready for ignition
state. Yes, it is a (somewhat mildly) a function of composition, but
at this temperature these models have very little evolution left before
the flame would be born even with differing composition. We have updated
the text (Section 2.2) to better explain this issue.

> - Why was the C+O model constructed with the same central temperature and
> density as the mixed hybrid model? Would it be more appropriate to follow
> a C+O white dwarf with MESA until some runaway condition was accomplished?

> - Related to the last point above: it is not clear how was assigned the
> composition profile to the C+O model. In particular, why is the abundance
> of 22Ne in the C+O model smaller than that in the hybrid mixed model? Does
> it has an impact on the final yield of 56Ni?

The referee raises an interesting point about how one might perform
these parameter studies. Indeed, a more realistic C/O progenitor as would
be produced by MESA would be "state of the art." The parameterized C/O
models we chose are motivated by results from stellar evolution and, in
particular, have compositions that reflect the effect of ``simmering",
that is slow convective burning, in the late stages of evolution as
the WD approaches the Chandrasekhar limit. This effect is observed as
the lower C fraction interior to about r =1.2e8 cm (the solid curve
in the bottom panel of Figure 1). Accordingly, while parameterized,
the structure of C/O progenitor does reflect our current understanding
of the state just prior to ignition.

Also, earlier studies (Jackson et al. 2010) indicated that the allowing
the metallicity to dictate a change in the DDT threshold density gave
a substantial change in the final 56Ni yield. We did not consider this
effect of metallicity by restricting the comparison here to one DDT
threshold density.

For this study we decided to use these parameterized C-O models because
being able to construct them to match the central density and temperature
of the hybrid models allowed us to better control the differences
between the hybrid and traditional models. More realistic C-O models
would not match the conditions of the hybrid models, and as we have
shown in previous work, differences in density can have a substantial
effect on the yields.

We have added text to sections 2.2 and 2.4 of the manuscript and
slightly re-arranged the description to better explain the motivation
for our choice.

> - Why were the models calculated in this paper ignited at the center,
> unlike the models in Willcox et al (2016)? Note that you invoke the
> off-center ignition as one of the causes of the differing results, but you
> do not provide a justification of the different ignition configuration.

Something else we could have better explained. The hybrid models
in Willcox et al have a very different structure with the highest
temperature, where ignition of the deflagration is expected, at a radius
of about 300 km.  The profiles are illustrated in Figures 2 and 9 of
Willcox et al. Yet another thing we have made clear in the revised
manuscript. In this case, we added text to section 2.4.

> - In the first paragraph of section 3, it is argued that the yield of 56Ni
> was approximated by assuming the composition upon NSE freeze-out is 56Ni
> plus equal parts 54Fe and 58Ni. However, in many models of supernova Ia
> there is a large abundance of 56Ni in the layers that do not go beyond
> QNSE. Have you corrected for this effect somehow?

The estimate of 56Ni yield does follow from assuming the NSE mix is
composed of 56Ni plus equal parts 54Fe and 58Ni, and this approach is
how all 56Ni in the simulation is calculated. 56Ni in (outer) layers
that did not go beyond QNSE would have been synthesized by fusion going
to the Fe-group and then mixed. As such, the estimate we use should be
robust. We have added a sentence to this paragraph of section 3 with
citations and an additional citation to previous work verifying this
approach in a variety of scenarios to the paragraph.

> *General*

> - Which is the role played by the hybrid progenitors in the global plot of
> supernovae type Ia? Since the lower limit of the 56Ni mass distribution
> is 0.6 solar masses, does it mean that hybrid progenitors lead to
> normal-to-bright supernovae while C+O progenitors represent the brighter
> end of this kind of supernovae? Does it imply a dominant contribution
> of double degenerate channel for the normal supernovae Ia? Most readers
> would thank a brief discussion of these and related issues.

These are the questions that we are exploring with this and related work,
and we regret not make these clear in the introductory material. We have
added text at the end of section 1.3 addressing exactly these issues.

> - Please, provide the values of the carbon abundance at the center and
> the mean of it in the white dwarf, in the initial model as well as in
> the final model of the MESA calculation.

The central carbon abundance in the MESA model is 0.1419 and the initial
model for Flash is constructed to have the same central carbon abundance.
The central carbon abundance in the C-O model is 0.4. We have added
these numbers to the manuscript. The mean values of C abundances were
not calculated with the models. We could integrate the models to obtain
a mass-weighted average, but we are not sure of the utility of this
quantity as our results show that the yields are a sensitive function
of the density. We prefer to just state the central C abundances and
present the profiles as we did in Figure 1 and hope this is acceptable.

> - Subsections 1.1 and 1.2 are more typical of a review than a regular
> paper, and are not at all necessary to understand what follows. They
> should be shortened to one or two medium length paragraphs. By the way,
> the reference Nomoto et al (1984) does not match the description of the
> flame that precedes it.

We have substantially shortened the sections, including removing the
Nomoto (1984) reference (good catch!). We modified both sections with an
emphasis on just the points important to our analysis (e.g.  the duration
of the deflagration phase and expansion).

> - Figures 3, 6, and 7 are not quite informative, and the same applies
> to the short mention of them in the text. The authors should consider
> removing them.

We appreciate the referee's point. The figures show the evolution of
quantities like 56Ni abundance with time and are similar to some of
the figures showing abundance evolution in Willcox et al. Because they
allow the reader to both visualize the scatter in the yields and also the
character of the burning (such as the DDT), we feel they are useful and
would like to keep them. Also, they are similar to figures in Willcox
et al, which allows comparison to results from another class of hybrid
progenitors. We agree with the referee that the discussion was scant
and have improved that.

In summary. we have taken the referee's criticism and suggestions and
revised the manuscript accordingly. The result is a significantly improved
version we trust is ready for publication.